Modeling Galaxy Spectral Energy Distributions in the SDSS Spectroscopic Survey

### 2002-02-05 docs/paper/paper.tex
fix flux units
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@17 75dff457-8751-0410-ab1a-cd701654ce6b
units
git-svn-id: svn+ssh://howdy.physics.nyu.edu/usr/local/svn/kcorrect/trunk@18 75dff457-8751-0410-ab1a-cd701654ce6b

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}}, and
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}
%David W. Hogg\altaffilmark{\ref{NYU}}, and
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

@@

The method described here for doing so follows closely the methods of
estimating photometric redshifts used by \citet{csabai00a} and
\citet{budavari00a}. It is designed to take advantage of what we
already understand about galaxy spectral energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.
Beginning with a space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), one can find an
orthonormal set of spectra spanning that space. To be more precise, we
define the dot-product in the space of galaxy spectra to be:
\begin{equation}
\vv{b}_i \cdot \vv{b}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
\lambda b_i(\lambda) b_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range corresponding to the range
which is within the observed wavelength range over most of the
redshift range of the sample. Otherwise, orthogonality is defined in
terms of regions of the spectra which are poorly constrained, which
make interpretation of the fits inconvenient.  This choice does not
affect the results for the reconstructed SED fit quantitatively
whatsoever.\footnote{If the template spectra were identical in the
subrange chosen and only differed outside of it, this choice would
cause computational problems, but in fact the spectra are all
independent in the restricted, as well as the full, wavelength range.}

@@

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& ZZZZ - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& ZZZZ - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons which are detected as a
function of wavelength (a unitless quantity).

@@

With a model of the rest-frame galaxy SED in hand, it is 
straightforward to project that SED onto a filter corresponding to 
an observation at any redshift, and predict the so-called
$K$-correction to the galaxy magnitude in order to compare galaxies 
at different redshifts. One simply calculates the AB magnitude at 
any redshift as:
\begin{equation}
m_{AB} = ZZZZ - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
f(\frac{\lambda}{1+z}) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].
\end{equation}
The simple way to think about this equation is that simply moving away
from the Earth doing does not change the bolometric flux of the
galaxy. So when the spectrum is stretched by the $(1+z)$ redshift
factor, as in $f(\lambda/(1+z))$, the bolometric flux should be
constant; that is, you must apply the extra multiplicative factor
$1/(1+z)$ to counteract the stretching of the spectrum. Once you have
have accounted for the fact that you are observing a particular
bandpass given the redshift of the object, you can then go ahead and
apply to the calculated magnitude the cosmological distance modulus
(that is, the luminosity distance-squared law, as tabulated by, {\it
e.g.} \citealt{hogg97a}).

>>>>

% Authorship determined by those I was directly involved with
% in performing this work as well as those responsible for the photo-z
% plates and those who have characterized the filter curves (since
% these might be publicly distributed). 
\author{
Michael R. Blanton\altaffilmark{\ref{NYU}},
%Tamas Budavari\altaffilmark{\ref{JHU}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
Istv\'an Csabai\altaffilmark{\ref{JHU}},
Mamoru Doi\altaffilmark{\ref{Tokyo}},
%Daniel Eisenstein\altaffilmark{\ref{Arizona}},
James E. Gunn\altaffilmark{\ref{Princeton}}, and
David W. Hogg\altaffilmark{\ref{NYU}}
%David J. Schlegel\altaffilmark{\ref{Princeton}}
%Julianne Dalcanton\altaffilmark{\ref{UW}},
%Jon Loveday\altaffilmark{\ref{Sussex}},
%Michael A. Strauss\altaffilmark{\ref{Princeton}},
%Mark SubbaRao\altaffilmark{\ref{Chicago}},
%David H. Weinberg\altaffilmark{\ref{Ohio}},
%John E. Anderson, Jr.\altaffilmark{\ref{Fermilab}},
%James Annis\altaffilmark{\ref{Fermilab}},
%Neta A. Bahcall\altaffilmark{\ref{Princeton}},
%Mariangela Bernardi\altaffilmark{\ref{Chicago}},
%J. Brinkmann\altaffilmark{\ref{APO}},
%Robert J. Brunner\altaffilmark{\ref{Caltech}},
%Scott Burles\altaffilmark{\ref{Fermilab}},
%Larry Carey\altaffilmark{\ref{UW}},
%Francisco J. Castander\altaffilmark{\ref{Chicago}, \ref{Pyrenees}},
%Andrew J. Connolly\altaffilmark{\ref{Pitt}},
%Istv\'an Csabai\altaffilmark{\ref{JHU}},
%Douglas Finkbeiner\altaffilmark{\ref{Berkeley}},
%Scott Friedman\altaffilmark{\ref{JHU}},
%Joshua A. Frieman\altaffilmark{\ref{Fermilab}},
%Masataka Fukugita\altaffilmark{\ref{CosmicRay},\ref{IAS}},
%G. S. Hennessy\altaffilmark{\ref{USNO}},
%Robert B. Hindsley\altaffilmark{\ref{USNO}},
%Takashi Ichikawa\altaffilmark{\ref{Tokyo}},
%\v{Z}eljko Ivezi\'{c}\altaffilmark{\ref{Princeton}},
%Stephen Kent\altaffilmark{\ref{Fermilab}},
%G. R.~Knapp\altaffilmark{\ref{Princeton}},
%D. Q.~Lamb\altaffilmark{\ref{Chicago}},
%R. French Leger\altaffilmark{\ref{UW}},
%Daniel C. Long\altaffilmark{\ref{APO}},
%Robert H. Lupton\altaffilmark{\ref{Princeton}},
%Timothy A.~McKay\altaffilmark{\ref{Michigan}},
%Avery Meiksin\altaffilmark{\ref{Edinburgh}},
%Aronne Merelli\altaffilmark{\ref{Caltech}},
%Jeffrey A. Munn\altaffilmark{\ref{USNO}},
%Vijay Narayanan\altaffilmark{\ref{Princeton}},
%Matt Newcomb\altaffilmark{\ref{CarnegieMellon}},
%R. C. Nichol\altaffilmark{\ref{CarnegieMellon}},
%Sadanori Okamura\altaffilmark{\ref{Tokyo}},
%Russell Owen\altaffilmark{\ref{UW}},
%Jeffrey R.~Pier\altaffilmark{\ref{USNO}},
%Adrian Pope\altaffilmark{\ref{JHU}},
%Marc Postman\altaffilmark{\ref{STScI}},
%Thomas Quinn\altaffilmark{\ref{UW}},
%Constance M. Rockosi\altaffilmark{\ref{Chicago}},
%Donald P. Schneider\altaffilmark{\ref{PennState}}, 
%Kazuhiro Shimasaku\altaffilmark{\ref{Tokyo}},
%Walter A. Siegmund\altaffilmark{\ref{UW}},
%Stephen Smee\altaffilmark{\ref{Maryland}},
%Yehuda Snir\altaffilmark{\ref{CarnegieMellon}},
%Chris Stoughton\altaffilmark{\ref{Fermilab}},
%Christopher Stubbs\altaffilmark{\ref{UW}},
%Alexander S.~Szalay\altaffilmark{\ref{JHU}},
%Gyula P.~Szokoly\altaffilmark{\ref{Potsdam}},
%Aniruddha R.~Thakar\altaffilmark{\ref{JHU}},
%Christy Tremonti\altaffilmark{\ref{JHU}},
%Douglas L. Tucker\altaffilmark{\ref{Fermilab}},
%Alan Uomoto\altaffilmark{\ref{JHU}},
%Dan vanden Berk\altaffilmark{\ref{Fermilab}},
%Michael S. Vogeley\altaffilmark{\ref{Drexel}},
%Patrick Waddell\altaffilmark{\ref{UW}},
%Brian Yanny\altaffilmark{\ref{Fermilab}},
%Naoki Yasuda\altaffilmark{\ref{NAOJ}},
%and Donald G.~York\altaffilmark{\ref{Chicago}}
}

@@

In the future, all observations of galaxies (and indeed of the sky in
general) will be performed using integral field spectrographs or
equivalent devices which combine high spatial resolution, high
spectral resolution, and high signal-to-noise. By that time, the
phenomenon of broad-band filters and the quaint nomenclature
surrounding their usage --- magnitudes, $K$-corrections, color terms,
{\it etc.} --- will have long since been forgotten. However, today
virtually all observations of galaxies are made through such filters,
and special care has to be taken to recover knowledge of galaxy
spectral energy distributions (SEDs) from these observations. This
paper describes how to do so in a way well-suited for recovering (for
example) fixed-frame magnitudes of objects over a large range of
redshifts.

@@

The method described here for doing so follows closely the methods of
estimating photometric redshifts used by \citet{csabai00a} and
\citet{budavari00a}. It is designed to take advantage of what we
already understand about galaxy spectral energy distributions.
%In addition, as implemented here, the method is
%equivalent to a fit to the star-formation history of the galaxies
%using a spectrophotometric model; we will investigate the implications
%of the model fits in a separate paper and concentrate here on the
%empirical aspects of the results.
Beginning with a space defined by $N_b$ template galaxy SEDs
$\vv{v}_i(\lambda)$ (say from \citealt{bruzual93a}), one can find an
orthonormal set of spectra spanning that space. To be more precise, we
define the dot-product in the space of galaxy spectra to be:
\begin{equation}
\vv{b}_i \cdot \vv{b}_j =
\int_{\lambda_{\mathrm{min}}}^{\lambda_{\mathrm{max}}} d\lambda
b_i(\lambda) b_j(\lambda),
\end{equation}
where $\lambda_{\mathrm{min}}$ and $\lambda_{\mathrm{max}}$ define the
wavelength range used to define orthogonality. These limits are
defined not over the full range covered by the $\vv{v}_i(\lambda)$
spectra, but instead over a smaller range corresponding to the range
which is within the observed wavelength range over most of the
redshift range of the sample. Otherwise, orthogonality is defined in
terms of regions of the spectra which are poorly constrained, which
make interpretation of the fits inconvenient.  This choice does not
affect the results for the reconstructed SED fit quantitatively
whatsoever.\footnote{If the template spectra were identical in the
subrange chosen and only differed outside of it, this choice would
cause computational problems, but in fact the spectra are all
independent in the restricted, as well as the full, wavelength range.}

@@

The magnitudes are calibrated $AB$, meaning that the system is
designed such that:
\begin{eqnarray}
m_{AB} &=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \lambda f(\lambda) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right]
\cr
&=& -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\nu \nu^{-1} f(\nu) R(\nu)}
{\int_{0}^{\infty} d\nu \nu^{-1} R(\nu)}\right],
\cr
\end{eqnarray}
where $R(\lambda)$ is the fraction of photons which are detected as a
function of wavelength (a unitless quantity).

@@

With a model of the rest-frame galaxy SED in hand, it is 
straightforward to project that SED onto a filter corresponding to 
an observation at any redshift, and predict the so-called
$K$-correction to the galaxy magnitude in order to compare galaxies 
at different redshifts. One simply calculates the AB magnitude at 
any redshift as:
\begin{equation}
m_{AB} = -48.60 - 2.5 \log_{10}\left[
\frac{\int_{0}^{\infty} d\lambda \frac{\lambda}{1+z}
f(\frac{\lambda}{1+z}) R(\lambda)}
{\int_{0}^{\infty} d\lambda \lambda^{-1} R(\lambda)}\right].
\end{equation}
The simple way to think about this equation is that simply moving away
from the Earth doing does not change the bolometric flux of the
galaxy. So when the spectrum is stretched by the $(1+z)$ redshift
factor, as in $f(\lambda/(1+z))$, the bolometric flux should be
constant; that is, you must apply the extra multiplicative factor
$1/(1+z)$ to counteract the stretching of the spectrum. Once you have
have accounted for the fact that you are observing a particular
bandpass given the redshift of the object, you can then go ahead and
apply to the calculated magnitude the cosmological distance modulus
(that is, the luminosity distance-squared law, as tabulated by, {\it
e.g.} \citealt{hogg97a}).
